---
title: "Buoy Assignment"
name: "Yuan Wang"
format: html
editor: visual
---

1.  Reading the buoy data.

```{r}
#Help from ChatGPT

# library(dplyr)
# library(readr)
# 
# # Years to download
# years <- 1984:2024
# 
# # Standard column names for StdMet
# col_names <- c("YY", "MM", "DD", "hh", "mm", "WDIR", "WSPD", "GST", "WVHT",
#                "DPD", "APD", "MWD", "PRES", "ATMP", "WTMP", "DEWP", "VIS", "PTDY", "TIDE")
# 
# # Base URL
# base_url <- "https://www.ndbc.noaa.gov/view_text_file.php?filename=44013h"
# 
# # Initialize list to store yearly data
# all_data <- list()
# 
# for (year in years) {
#   url <- paste0(base_url, year, ".txt.gz&dir=data/historical/stdmet/")
#   local_file <- paste0("44013h", year, ".txt.gz")
#   
#   # Download file
#   download.file(url, destfile = local_file, mode = "wb")
#   
#   # Read file: space-delimited, skip empty lines
#   df <- read_table(local_file, col_names = FALSE, na = c("MM", "99", "99.0", "99.00", "999", "9999", "999.0"))
#   
#   # Fill missing columns with NA
#   if (ncol(df) < length(col_names)) {
#     for (i in (ncol(df)+1):length(col_names)) df[[i]] <- NA
#   }
#   
#   # Assign standard column names
#   colnames(df) <- col_names[1:ncol(df)]
#   
#   all_data[[as.character(year)]] <- df
# }
# 
# # Combine all years
# combined_data <- bind_rows(all_data)
# 
# # Save as CSV
# write.csv(combined_data, "buoy_44013_stdmet_1984_2024.csv", row.names = FALSE)
# 
# cat("✅ CSV file created: buoy_44013_stdmet_1984_2024.csv\n")
```

Data column descriptions: https://www.ndbc.noaa.gov/faq/measdes.shtml

```{r}
df <- read.csv("buoy_44013_stdmet_1984_2024.csv", header = TRUE)
buoy_df <- df[,1:19]
head(buoy_df)

colnames(buoy_df) <- c(
  "YY","MM","DD","hh","WD","WSPD","GST","WVHT","DPD",
  "APD","MWD","BAR","ATMP","WTMP","DEWP","VIS", "PTDY","TIDE")
head(buoy_df)
```

2.  Address a question, make a plot, and explain your findings through it.

Question: What was the median wind speed in 2001?

```{r}
buoy_df$WSPD <- as.numeric(buoy_df$WSPD)

df_01 <- buoy_df[,c("YY", "WSPD")]
df_02 <- df_01 %>% 
  filter(YY == 2001)

boxplot(df_02$WSPD, 
        main = "Wind Speed in 2001",
        ylab = "Wind Speed")
```

Explanation: The box-plot shows that the median wind speed in 2001 was 5 m/s. The first quartile is at 3 m/s, and the third quartile is at 7 m/s. This means that 50% of wind speed measurements fall between 3 m/s and 7 m/s. We can also observe that the upper half of the distribution contains more variability, and there are several outliers representing unusually high wind speeds.
